---
title: "Homework 2: Individual"
author: "Michael Mendoza"
format: html
editor: visual
date: 31AUG2026
purpose: create HW2 content (individual)
---

# Question 1

The following code creates the pre-/post-treatment records as 2 different vectors

```{r}

pre_bp <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)
print(pre_bp)

post_bp <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)
print(post_bp)

```

The following Code assigns names to these vectors and creates a vector of these names.

```{r}
names(pre_bp) <- paste("Subject", 1:20, sep = "_")
print(pre_bp)

names(post_bp) <- paste("Subject", 1:20, sep = "_")
print(post_bp)

patient <- paste("Subject", 1:20, sep = "_")
print(patient)
```

The following stores the difference of the pre-/post- treatment records, a positive record indicates a decrease in blood pressure.

```{r}
diff_bp <- pre_bp - post_bp
print(diff_bp)
```

This is the mean of the difference of among all patients.

```{r}

print(mean(diff_bp))
```

The following returns a vector of which subjects experienced a positive change, and stores the records as a subset.

```{r}

print(which(diff_bp>0))

pos_subset <- diff_bp[which(diff_bp>0)]
```

This is the mean of the difference of among patients who experienced a decrease in blood pressure.

```{r}
print(mean(diff_bp))
```

# Question 2

The following code creates a data frame with 4 variables: patient, pre_bp, post_bp, and diff_bp

```{r}
bp_df <- data.frame(patient, pre_bp, post_bp, diff_bp)
```

Return only negative difference patients.

```{r}
print(bp_df[bp_df$diff_bp < 0, ])
```

Adding True/False flag to dataset when post_bp \< 120

```{r}
bp_df$lt_120 <- bp_df$diff_bp < 0
print(bp_df)
```

Knit code to print bp_df to final document.

```{r}
knitr::kable(bp_df, caption = 'Blood Pressure Data Table')
```

# Question 3

Create the new dataset.

```{r}
pre_bp2 <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)

post_bp2 <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)

names(pre_bp2) <- paste("Subject", 1:10, sep = "_")

names(post_bp2) <- paste("Subject", 1:10, sep = "_")

patient2 <- paste("Subject", 1:10, sep = "_")

diff_bp2 <- pre_bp2 - post_bp2

bp_df2 <- data.frame("patient" = patient2, "pre_bp" = pre_bp2, "post_bp" = post_bp2, "diff_bp" = diff_bp2)

bp_df2$lt_120 <- bp_df2$diff_bp < 0
```

Create the list with 2 elements: treatment, placebo.

```{r}
bp_list <- list("treatment" = bp_df, "placebo" = bp_df2)
```

3 different types of syntax to access the first list element.

```{r}

bp_list[[1]]

bp_list["treatment"]

head(bp_list, 1)

```

Access Placebo: pre_bp in a single line.

```{r}
print(bp_list$placebo$pre_bp)
```

# Question 4

Create new column named status in each dataset in listing

```{r}
bp_list$treatment$status <- character(20)

bp_list$placebo$status <- character(10)
```

Using if/else lolgic to derive status flag in treatment dataframe.

```{r}
for (i in 1:20) {
  
  if (bp_list$treatment$post_bp[i] <= 120) {
    bp_list$treatment$status[i] <- "optimal"
  }
    else if (bp_list$treatment$post_bp[i] > 130) {
      bp_list$treatment$status[i] <- "high"
    }
    else {
      bp_list$treatment$status[i] <- "borderline"
    }
}

print(bp_list$treatment)
```

Repeat for placebo data

```{r}
for (i in 1:10) {
  
  if (bp_list$placebo$post_bp[i] <= 120) {
    bp_list$placebo$status[i] <- "optimal"
  }
  else if (bp_list$placebo$post_bp[i] > 130) {
    bp_list$placebo$status[i] <- "high"
  }
  else {
    bp_list$placebo$status[i] <- "borderline"
  }
}

print(bp_list$placebo)
```
